Analysis

### 2023-03-30 Analysis.rmd
merge figures with number of forecasters with same figure for experts

# Table 1: Performance
================================================================================

@@

plot_number_forecasters <- avail |>
  pivot_wider(names_from = type, values_from = `Number forecasts`) |>
  mutate(Rt = Direct + Rt) |>
  pivot_longer(cols = c(Direct, Rt), names_to = "type",
               values_to = "Number forecasts") |>
  as.data.table() |>
  rbind(avail_ensemble, fill = TRUE) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             group = group,
             fill = group, 
             alpha = type)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  scale_fill_manual(values = colors[1:2]) +
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

ggsave("output/figures/num-forecasters.png", plot = plot_number_forecasters, 
       width = 7, height = 3.5)

@@

p_num_experts <- nums |>
  mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
  group_by(target_type, forecast_date, expert, method) |>
  summarise(n = n()) |>
  pivot_wider(names_from = method, values_from = n) |>
  mutate(Rt = Rt + direct) |>
  pivot_longer(cols = c(Rt, direct), names_to = "method", values_to = "n") |>
  ggplot(aes(x = forecast_date, y = n, 
             group = expert,
             fill = expert, 
             alpha = method)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

ggsave("output/figures/num-forecasters-expert.png", plot = p_num_experts, 
       width = 7, height = 3.5)

@@

plot_ranks <- function(scores, hor = 2) {
  scores |>
    summarise_scores() |>
    select(model, target_type, horizon, forecast_date, interval_score, scale, type_and_scale) |>
    group_by(target_type, scale, horizon, forecast_date) |>
    mutate(rank = rank(interval_score)) |>
    filter(horizon == hor) |>
    ggplot(aes(y = model, group = rank, fill = rank)) +
    geom_bar(position = position_stack(), colour = "white", size = 0.03) + 
    facet_wrap(~ type_and_scale) + 
    scale_fill_continuous(trans='reverse') + 
    theme_scoringutils() + 
    labs(y = "", x = "Frequency", fill = "Rank") + 
    theme(legend.position = "right")
}

p_ranks <- scores |>
  filter(model %in% models_non_human) |>
  mutate(model = factor(model, levels = rev(levels(models_non_human)))) |>
  plot_ranks()
ggsave("output/figures/performance-ranks.png", 
       plot = p_ranks, 
       width = 7, height = 3.4)

p_ranks_4 <- scores |>
  filter(model %in% models_non_human) |>
  mutate(model = factor(model, levels = rev(levels(models_non_human)))) |>
  plot_ranks(hor = 4)
ggsave("output/figures/performance-ranks-4.png", 
       plot = p_ranks_4, 
       width = 7, height = 3.4)

>>>>

# Table 1: Performance
================================================================================
- also includes Table SI.1

@@

@@

plot_number_forecasters <- avail |>
  pivot_wider(names_from = type, values_from = `Number forecasts`) |>
  mutate(Rt = Direct + Rt) |>
  pivot_longer(cols = c(Direct, Rt), names_to = "type",
               values_to = "Number forecasts") |>
  as.data.table() |>
  rbind(avail_ensemble, fill = TRUE) |>
  ggplot(aes(x = forecast_date, y = `Number forecasts`, 
             group = group,
             fill = group, 
             alpha = type)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  scale_fill_manual(values = c(
    "Expert" = "#984EA3", 
    "Non-Expert" = "#377EB8", 
    "Forecast Hub" = "#F8766D", 
    "Human forecast" = "#7CAE00"), 
    breaks = c("Expert", "Non-Expert", "Forecast Hub", "Human forecast")
    ) +
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin())

p_num_experts <- nums |>
  mutate(expert = ifelse(expert, "Expert", "Non-Expert")) |>
  group_by(target_type, forecast_date, expert, method) |>
  mutate(method = ifelse(method == "direct", "Direct", "Rt")) |>
  summarise(n = n()) |>
  pivot_wider(names_from = method, values_from = n) |>
  mutate(Rt = Rt + Direct) |>
  pivot_longer(cols = c(Rt, Direct), names_to = "method", values_to = "n") |>
  ggplot(aes(x = forecast_date, y = n, 
             group = expert,
             fill = expert, 
             alpha = method)) +
  geom_bar(stat = "identity", 
           color = "white",
           linewidth = 0.2,
           width = 5,
           position = position_dodge(width = 5)
           ) + 
  facet_wrap(~ target_type) + 
  scale_alpha_manual(values = c(1, 0.5), na.translate = FALSE) + 
  scale_fill_manual(values = c(
    "Expert" = "#984EA3", 
    "Non-Expert" = "#377EB8", 
    "Forecast Hub" = "#F8766D", 
    "Human forecast" = "#7CAE00"), 
    breaks = c("Expert", "Non-Expert", "Forecast Hub", "Human forecast")
    ) +
  labs(x = "Forecast date", fill = "Forecast type", 
       alpha = "Forecast method") + 
  theme_scoringutils() +
  theme(legend.position = "bottom", 
        legend.box="vertical", legend.margin = margin()) + 
  labs(y = "Number forecasts")

p_nums <- plot_number_forecasters /
  p_num_experts +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "right")

ggsave("output/figures/num-forecasters.png", plot = p_nums, 
       width = 7, height = 5)

@@

ranks <- scores |>
  filter(model %in% models_non_human) |>
  summarise_scores() |>
  select(model, target_type, horizon, forecast_date, interval_score, scale, type_and_scale) |>
  group_by(target_type, scale, horizon, forecast_date) |>
  mutate(rank = rank(interval_score))

plot_ranks <- function(ranks, hor = 2) {
  ranks |>
    filter(horizon == hor) |>
    ggplot(aes(y = model, group = rank, fill = rank)) +
    geom_bar(position = position_stack(), colour = "white", size = 0.03) + 
    facet_wrap(~ type_and_scale) + 
    scale_fill_continuous(trans='reverse') + 
    theme_scoringutils() + 
    labs(y = "", x = "Frequency", fill = "Rank") + 
    theme(legend.position = "right")
}

p_ranks <- ranks |>
  mutate(model = factor(model, levels = rev(levels(models_non_human)))) |>
  plot_ranks()
ggsave("output/figures/performance-ranks.png", 
       plot = p_ranks, 
       width = 7, height = 3.4)

p_ranks_4 <- ranks |>
  mutate(model = factor(model, levels = rev(levels(models_non_human)))) |>
  plot_ranks(hor = 4)
ggsave("output/figures/performance-ranks-4.png", 
       plot = p_ranks_4, 
       width = 7, height = 3.4)

ranks |>
  filter(horizon == 2) |>
  mutate(model = factor(model, levels = (levels(models_non_human)))) |>
  group_by(model, target_type, scale, rank) |>
  summarise(n = n()) |>
  mutate(rel = n() / 13) |>
  pivot_wider(names_from = rank, values_from = c(n, rel)) |>
  {\(.) {replace(.,is.na(.),0)}}() |>
  mutate(n_1 = paste0(n_1, " (", round(rel_1, 2), ")"), 
         n_2 = paste0(n_2, " (", round(rel_2, 2), ")"), 
         n_3 = paste0(n_3, " (", round(rel_3, 2), ")"), 
         n_4 = paste0(n_4, " (", round(rel_4, 2), ")")) |>
  select(-c(rel_1, rel_2, rel_3, rel_4))
